---
title: ""
author: ""
date: ""
output: html_document
---
## **Introduction to building R Packages**


## **1. Package creation**

### A package about the ordinary non-parametric bootstrap

You are requested to write a `R` package (`footstrap`) that allows users to carry out a bootstrap analysis on the `t` statistic to test for equal means. The package should focus on the ordinary non-parametric bootstrapping procedure (i.e. the very basic one). It will also compare the bootstrap results with the theoretical ones (approximation of the statistic distribution to a `Student t` with specific degrees of freedom).

### Functions inside the package

The programmer needs to provide:

  1. The _statistic_ that the main function will need to run on each bootstrap sample: `get.lm`
  2. the _main function_ `get.bootstrap`, that runs the bootstrap simulation (i.e., it uses `get.lm`)
  3. additional functions `summary.get.bootstrap` and `plot.get.bootstrap` in order to obtain a summary of the results and plots
 
The following diagram illustrates the structure of the functions.

  `put here the flow diagram`
  

### Let's get our hands dirty with coding!

#### 1. **_statistic_**: `get.lm`

```{r}
get.lm <- function(formula, data, indices) {
  return(lm(formula = formula, data = data[indices, ])$coef)
}
```

#### 2. **_the main function_**: `get.bootstrap`

```{r}
get.bootstrap <- function(data, statistic, replicates, formula) {
    # Get the number of rows.
    n <- nrow(data)

    # Create random samples with replacement.
    samples <- matrix(sample(1:n, n * replicates, replace = T), nrow = replicates, ncol = n)

    # Run the bootstrap over the samples.
    boot.sim <- t(apply(samples, 1, function(x) statistic(formula = formula, data = data, indices = x)))

    # Create the output object.
    out <- list()
    out$boot.sim <- boot.sim
    out$obs.statistic <- statistic(formula = formula, data = data, indices = 1:n)
    out$data <- data
    out$formula <- formula

    # Define object class.
    class(out) <- "boot"

    return(out)
}
```

#### 3. **_summary of the results_**: `summary.boot`

```{r}
summary.boot <- function(boot) {
    # MLE standard errors.
    se.mle <- summary(lm(formula = boot$formula, data = boot$data))$coef[, 2]

    # Bootstrap standard errors.
    se.boot <- apply(boot$boot.sim, 2, sd)

    # Put all SEs together.
    se.all <- cbind(se.mle, se.boot)

    # Add column names.
    colnames(se.all) <- c("MLE", "Bootstrap")

    # Pretty print.
    cat("Comparison between Standard Errors:", "\n")
    print(se.all)
}
```

#### 4. **_plotting of the results_**: `plot.boot`

```{r}
plot.boot <- function(boot, breaks = 50) {
    # This line is only inteded for the handout!
    require(grDevices)
    
    # Get number of parameters.
    n.parameters <- dim(boot$boot.sim)[2]

    # Split screen.
    par(mfrow = rev(n2mfrow(n.parameters)))

    # Create plot template.
    for(i in 1:n.parameters) {
        hist(boot$boot.sim[, i], breaks = breaks, 
             freq = FALSE, 
             xlab = colnames(boot$boot.sim)[i], 
             main = colnames(boot$boot.sim)[i], 
             col = "papayawhip")
        box()
        abline(v = boot$obs.statistic[i], col = "red", lwd = 2, lty = 2)
        legend("topright", legend = c("MLE"), col = c("red"), lwd = 2, 
               lty = 2, ncol = 1, cex = 0.8, bty = "n")
    }

    mtext("Bootstrap distributions", side = 3, line = -1, outer = TRUE)

    # Un-split screen.
    par(mfrow = c(1, 1))
}
```


---


## **2. Package testing**

Let's load our new package `footstrap` and run the examples in the codes below.

```{r, message=FALSE,warning=FALSE}
# library(footstrap)
```

#### **Example 1** model = `speed ~ dist`, dataset `cars` (sample size = 50)
```{r, message=FALSE,warning=FALSE}
library(datasets)
head(cars)
example.1 <- get.bootstrap(data = cars, statistic = get.lm, replicates = 1e3, formula = speed ~ dist)
plot(example.1)
summary(example.1)
```

#### **Example 2** model = `weight ~ feed`, dataset `chickwts` (sample size = 71)
```{r, message=FALSE,warning=FALSE}
head(chickwts)
example.2 <- get.bootstrap(data = chickwts, statistic = get.lm, replicates = 1e3, formula = weight ~ feed)
plot(example.2)
summary(example.2)
```

#### **Example 3** where `log(Volume) ~ log(Girth)`, dataset `trees` (sample size = 31)
```{r, message=FALSE,warning=FALSE}
head(trees)
example.3 <- get.bootstrap(data = trees, statistic = get.lm, replicates = 1e3, formula = log(Volume) ~ log(Girth))
plot(example.3)
summary(example.3)
```
