---
title: "Introduction to building R Packages"
author: ""
date: ""
output: html_document
---

We decided to estimate a linear regression model by maximizing its loglikelihood function.

## The data frame
First of all, let's have a look into the data structure, which our regressors are and which outcome we are interested in.

```{r, message=FALSE,warning=FALSE}
library(MASS)
head(crabs)
```
> **Description of the data frame** (_from the documentation_)  
> The `crabs` data frame has 200 rows and 8 columns, describing 5 morphological measurements on 50 crabs each of two colour forms and both sexes, of the species _Leptograpsus variegatus_ collected at Fremantle, W. Australia.


### The statistical unit
The _statistical unit_ in the data frame is a _crab_. For each of the 200 crabs several morphological measurements were recorded, such as

* `FL`  : _Front Lobal Size_ (mm);
* `RW`  : _Rear Width_ (mm);
* `CL`  : _Carapace Length_ (mm);
* `CW`  : _Carapace Width_ (mm);
* `BD`  : _Body Depth_ (mm).

Other than these, we also know

* `sex` : `M` (_Male_), `F` (_Female_); 
* `sp`  : _specie_, referring to the color forms, `B` for _blue_, `O` for _orange_.

Finally, the `index` column assumes values in `1:50` within each of the four groups that result by the combination of the levels of `sex` and `sp` (specie). We won't use it in our analysis.

### Goal of the research
Both color forms had been classified as morphs of the _L. variegatus_. However, with genetic analysis and ecological studies, researchers found these two forms to be two distinct species. 
Therefore, the need of developing a criteria for the identification of the specie.

## The model: Logistic Regression
Let's consider `sp`as our _dependent variable_ $Y$. We want to model the probability of observing a _blue_ crab (that is $P(Y = B)$) given the morphological measurements in our hands, that we will call $X$. We assume that
$$Y_i\sim Ber(p_i)\quad{}i=1,\ldots,200$$
$Y_i$ generates from a _Bernoulli_ distribution with probability of assuming the value $B$
$$P(Y_i = B) = p_i$$
We would like to model this probability as a function of the morphological measurements listed above. We can model $p_i$ as a function of them
$$p_i = f(X_i) $$
without overthinking about it, we can just use the most used (on average) function that is the **logistic function**,
$$f(X_i) = \frac{\exp{\left\lbrace\eta_i\right\rbrace}}{1+\exp{\left\lbrace\eta_i\right\rbrace}}$$
where $\eta_i$ is the **linear predictor** 
$$\eta_i = \beta_0 + \beta_1\cdot sex^{male}_i + \beta_2\cdot FL_i + \beta_3 \cdot RW_i + \beta_4 \cdot CL_i + \beta_5 \cdot CW_i + \beta_6 \cdot BD_i $$
Parameters are interpreted in the usual fashion: they represent the effect of a specific variable on the log-odds ratio.
# The likelihood to maximize
Write here the likelihood (independence assumption among observation)
Write here the function we need to minimize (-loglikelihood)

Task : Researchers want statistician to write a package with all the functions needed for the analysis

# Functions
* model matrix
* likelihood function
* optimization function
* printing results
* plotting results

```{r}

```